Oncogenomics Report for patient ID `r patientID`
========================================================

<!-- dataTables -->
<script src="http://ajax.aspnetcdn.com/ajax/jquery.dataTables/1.9.4/jquery.dataTables.js"></script>
<link rel="stylesheet" href="http://ajax.aspnetcdn.com/ajax/jquery.dataTables/1.9.4/css/jquery.dataTables.css"/>
<script src="http://next.datatables.net/release-datatables/examples/resources/bootstrap/3/dataTables.bootstrap.js"></script>
<link rel="stylesheet" href="http://next.datatables.net/release-datatables/examples/resources/bootstrap/3/dataTables.bootstrap.css"/>

<!-- TableTools -->
<script src="http://cdn.aldu.net/jquery.datatables/extras/tabletools/2.0.3/js/TableTools.js"></script>
<link rel="stylesheet" href="http://cdn.aldu.net/jquery.datatables/extras/tabletools/2.0.3/css/TableTools.css"/>

<!-- ZeroClipboard -->
<script src="http://cdn.aldu.net/jquery.datatables/extras/tabletools/2.0.3/js/ZeroClipboard.js"></script>

<!-- ColVis -->
<script src="http://cdn.eventcore.com/js/plugins/dataTables.ColVis-1.07.js"></script>

<!-- Scientific notation sorting -->
<script src="../media/src/scientific.js"></script>

<!-- read in external core code -->
```{r external_code, echo=FALSE}
  read_chunk(paste(config$R_SOURCE_PATH, '/src/core/setup.R', sep=''))
  read_chunk(paste(config$R_SOURCE_PATH, '/src/core/dataIn.R', sep=''))
  read_chunk(paste(config$R_SOURCE_PATH, '/src/core/gene_table.R', sep=''))
```

<!-- read in all optional modules from the module folder -->
```{r module_code, echo=FALSE}
  read_chunk(paste(config$R_SOURCE_PATH, '/src/modules/qual_ctrl.R', sep=''))
  read_chunk(paste(config$R_SOURCE_PATH, '/src/modules/reliable_gep.R', sep=''))
  read_chunk(paste(config$R_SOURCE_PATH, '/src/modules/diff_exp.R', sep=''))
  read_chunk(paste(config$R_SOURCE_PATH, '/src/modules/brca_mol_class.R', sep=''))
  read_chunk(paste(config$R_SOURCE_PATH, '/src/modules/spia.R', sep=''))
  read_chunk(paste(config$R_SOURCE_PATH, '/src/modules/PW_tables_and_viz.R', sep=''))
  read_chunk(paste(config$R_SOURCE_PATH, '/src/modules/fusion_catcher.R', sep=''))
  read_chunk(paste(config$R_SOURCE_PATH, '/src/modules/intogen.R', sep=''))
  read_chunk(paste(config$R_SOURCE_PATH, '/src/modules/dog.R', sep=''))
  read_chunk(paste(config$R_SOURCE_PATH, '/src/modules/snpAnnotation.R', sep=''))
```

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
```

```{r data, echo=FALSE}
```

# Quality Control & Preprocessing

## Raw data QC

The quality of the RNA-seq reads was assessed using FASTQC `r citep("http://www.bioinformatics.babraham.ac.uk/projects/fastqc")`. The table below lists basic QC statistics. Click on the provided links to open up the full FastQC report.


```{r fastQC, echo=FALSE, results='asis'}
```

******

## Alignment

Reads were aligned to the human genome (hg19) using the STAR aligner `r citep("10.1093/bioinformatics/bts635")`. Resulting .SAM files were sorted and converted to .BAM files using SAMtools software `r citep("10.1093/bioinformatics/btp352")`. The table below lists the alignment statistics as provided by the STAR output:

<div style="width:40%;padding:0 10pt 0 0;float:left;">
```{r starQCTable, echo=FALSE, results='asis', warning=FALSE}
```
</div>

<div style="width:60%;padding:0 10pt 0 0;float:right;">
```{r starQCPie, echo=FALSE, results='asis'}
```
</div>
<div style="clear:both;"></div>

## Insert size distribution

Insert size distribution is calculated using Picard tools [CollectInsertSizeMetrics](http://picard.sourceforge.net/command-line-overview.shtml#CollectInsertSizeMetrics) function `r citep("http://picard.sourceforge.net/")`. The table lists statistics as provided by this tool:

<div style="width:40%;padding:0 10pt 0 0;float:left;">
```{r InsertSizeTable, echo=FALSE, results='asis', warning=FALSE}
```
</div>

<div style="width:60%;padding:0 10pt 0 0;float:right;">
```{r InsertSizeImage, echo=FALSE, results='asis'}
```
</div>
<div style="clear:both;"></div>

## RNA sequencing metrics

Metrics are assessed using Picard tools [CollectRnaSeqMetrics](http://picard.sourceforge.net/command-line-overview.shtml#CollectRnaSeqMetrics) function `r citep("http://picard.sourceforge.net/")`. Metrics about the alignment of RNA to various functional classes of loci in the genome are collected: coding, intronic, UTR, intergenic, ribosomal. The table below lists the metrics provided by this tool:

<div style="width:40%;padding:0 10pt 0 0;float:left;">
```{r RNASeqMetricsTable, echo=FALSE, results='asis', warning=FALSE}
```
</div>

<div style="width:60%;padding:0 10pt 0 0;float:right;">
```{r RNASeqMetricsPie, echo=FALSE, warning=FALSE, results='asis'}
```
</div>
<div style="clear:both;"></div>

******

## Gene expression quantification 

Gene expression quantification was done using [htseq-count](http://www-huber.embl.de/users/anders/HTSeq/doc/count.html) function within the python HTSeq analysis package `r citep("http://www-huber.embl.de/users/anders/HTSeq/doc/overview.html")`, which counts all reads overlapping known exons.

<div style="width:40%;padding:0 10pt 0 0;float:left;">
```{r htseqQCTable, echo=FALSE, results='asis', warning=FALSE}
```
</div>

<div style="width:60%;padding:0 10pt 0 0;float:right;">
```{r htseqQCPie, echo=FALSE, results='asis'}
```
</div>
<div style="clear:both;"></div>

## Gene expression reliability

```{r reliableGEP, echo=FALSE, results='hide'}
```

Gene expresssion values are grouped into "reliable expressed genes" and "not so reliable expressed genes" based on a background distribution of `r length(patientREC$background)` genes which have a median expression value of 0 in the `r dim(reference)[2]` TCGA reference cohort:

* **reliable expressed** Number of genes reliable expressed in the sample
* **marginal expressed** Number of genes that are marginally (between 0.01 & 0.02) reliable expressed in the sample
* **not reliable expressed** Number of genes that are not reliable expressed in the sample

```{r reliableGEPTable, echo=FALSE}
```

```{r reliableGEPPlot, echo=FALSE, warning=FALSE, results='asis'}
```

******

Expression density plot

```{r reliableGEPDensityPlot, echo=FALSE, fig.width=20, fig.height=10, fig.align='center'}
```

******

# Analysis

## Differential Gene Expression

Differential gene expression is calculated based on a model using the negative binomial distribution as implemented in the DESeq2 package `r citep("10.1186/gb-2010-11-10-r106")`. 

<!-- run the differential expression on the patient samples -->
```{r diffExpressionNBinom, echo=FALSE, message=FALSE, eval=TRUE}
```

```{r diffExpPlot, echo=FALSE, warning=FALSE, results='asis'}
```

******

## Gene Expression Table

<!-- create gene expression table including results from diff. expression calling if available -->
* **Gene** HUGO Symbol for the gene 
* **Diff_expressed** [upregulated/downregulated] above/below median expression value within the normal control +/- three standard deviations 
* **Rank** gene rank within differentially expressed genes
* **perRank** percentile rank within the distribution of all genes within the patient sample
* **Patient** Deseq scaled and log2 transformed read count value of the patient sample
* **Normal** Deseq scaled and log2 transformed median read count value of the normal reference cohort (n=`r ifelse(any(group=='Healthy'), length(which(group=='Healthy')), length(which(group=='Normal')))`)
* **Normal_sd** standard deviation for each scaled read count value of the normal reference cohort
* **d_normal** delta between patient sample read count value and median read count value of the normal reference cohort
* **Ref** Deseq scaled and log2 transformed median read count value of the TCGA reference cohort (n=`r length(which(group==config$TUMOR_TYPE))`)
* **perRankRef** percentile rank within the distribution of all genes within the reference cohort
* **Ref_sd** standard deviation for each scaled read count value of the reference cohort
* **d_ref** delta between patient sample read count value and median read count value of the reference cohort
* **List_Gene** genes within the supplied gene list

```{r GeneTable, echo=FALSE, results='asis', message=FALSE, eval=TRUE}
```

<table id="gep_table">
    <thead>
        <tr>
            <th>Gene</th>
            <th>Diff_expressed</th>
            <th>Reliable_expressed</th>
            <th>Rank</th>
            <th>perRank</th>
            <th>Patient</th>
            <th>Normal</th>
            <th>Normal_sd</th>
            <th>d_normal</th>
            <th>Ref</th>
            <th>perRankRef</th>
            <th>Ref_sd</th>
            <th>d_ref</th>
            <th>Pvalue</th>
            <th>List_Gene</th>
            <th>GeneWiki</th>
        </tr>
    </thead>
    <tbody></tbody>
</table>

<script type="text/javascript" charset="utf-8">
    $(document).ready(function() {
    var table = $('#gep_table').dataTable({
      "sDom": 'T<"clear">lfrtip',
      "oTableTools": {
          "sSwfPath": "../media/swf/copy_csv_xls_pdf.swf",
          "aButtons": [
                "copy",
                "csv"
            ]
      },
      "aoColumns": [ 
      null,
      null,
      null,
      {"bVisible": false},
		  {"bVisible": false},
      null,
      null,
      {"bVisible": false},
      null,
      {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"sType": "scientific"},
      {"bVisible": false},
      null
		], 
    "iDisplayLength": 10,
    "bPaginate": true,
    "bProcessing": true,
    "sAjaxSource": 'json/gepall.txt'
    //"sScrollXInner": "110%",
    });
    });    
</script>

******

## Prediction of Receptor Status

```{r brcaMolClassData, echo=FALSE, message=FALSE}
```

* **Receptor** Receptor type
* **Status** predicted receptor status for each receptor type
* **Prediction_probability** prediction probability for each receptor type

```{r brcaReceptorStatus, echo=FALSE, results='asis'}
```

<script type="text/javascript" charset="utf-8">
    $(document).ready( function () {
    $('#receptor_table').dataTable({
          "sDom": 'T<"clear">lfrtip',
          "oTableTools": {
          "sSwfPath": "../media/swf/copy_csv_xls_pdf.swf",
          "aButtons": [
                "copy",
                "csv"
                ]
          }
    });
    });
</script>

******

## Prediction of molecular subtype

* **Subtype** prediction of one of the subtypes [Luminal A / Luminal B / Basal / Her2]
* **Prediction_probability** prediction probability for each receptor type

```{r brcaMolClass, echo=FALSE, results='asis'}
```

<script type="text/javascript" charset="utf-8">
    $(document).ready( function () {
    $('#mol_table').dataTable({
          "sDom": 'T<"clear">lfrtip',
          "oTableTools": {
          "sSwfPath": "../media/swf/copy_csv_xls_pdf.swf",
          "aButtons": [
                "copy",
                "csv"
                ]
          }
    });
    });
</script>

******

## Pathways

```{r spiaPW, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE}
```

```{r PWFunctions, echo=FALSE, message=FALSE, warning=FALSE}
```

### KEGG

```{r KEGGTable, message=FALSE, echo=FALSE, results='asis', eval=TRUE}
```

<script type="text/javascript" charset="utf-8">
    $(document).ready( function () {
    $('#kegg_table').dataTable({
          "sDom": 'T<"clear">lfrtip',
          "oTableTools": {
          "sSwfPath": "../media/swf/copy_csv_xls_pdf.swf",
          "aButtons": [
                "copy",
                "csv"
                ]
          },
      "aoColumns": [ 
      null,
			{"bVisible": false},
      {"bVisible": false},
		  {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"sType": "scientific"},
      {"bVisible": false},
      null,
      null
		]
    });
    });
</script>

******

### Biocarta

```{r BiocartaPlots, echo=FALSE, fig.show='hide', fig.width=20, fig.height=10, message=FALSE, results='hide', eval=TRUE}
```

```{r BiocartaTable, echo=FALSE, message=FALSE, results='asis', eval=TRUE}
```

<script type="text/javascript" charset="utf-8">
    $(document).ready( function () {
    $('#biocarta_table').dataTable({
          "sDom": 'T<"clear">lfrtip',
          "oTableTools": {
          "sSwfPath": "../media/swf/copy_csv_xls_pdf.swf",
          "aButtons": [
                "copy",
                "csv"
                ]
          },
      "aoColumns": [ 
      null,
  		{"bVisible": false},
      {"bVisible": false},
		  {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"sType": "scientific"},
      {"bVisible": false},
      null,
      null
		]        
    });
    });
</script>

******

### NCI

```{r NCIPlots, echo=FALSE, fig.show='hide', fig.width=20, fig.height=10, message=FALSE, results='hide', eval=TRUE}
```

```{r NCITable, echo=FALSE, message=FALSE, results='asis', eval=TRUE}
```

<script type="text/javascript" charset="utf-8">
    $(document).ready( function () {
    $('#nci_table').dataTable({
          "sDom": 'T<"clear">lfrtip',
          "oTableTools": {
          "sSwfPath": "../media/swf/copy_csv_xls_pdf.swf",
          "aButtons": [
                "copy",
                "csv"
                ]
          },
      "aoColumns": [ 
      null,
  		{"bVisible": false},
      {"bVisible": false},
		  {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"sType": "scientific"},
      {"bVisible": false},
      null,
      null
		]        
    });
    });
</script>

******

### Reactome

```{r ReactomePlots, echo=FALSE, fig.show='hide', fig.width=20, fig.height=10, message=FALSE, results='hide', eval=TRUE}
```

```{r ReactomeTable, echo=FALSE, message=FALSE, results='asis', eval=TRUE}
```

<script type="text/javascript" charset="utf-8">
    $(document).ready( function () {
    $('#reactome_table').dataTable({
          "sDom": 'T<"clear">lfrtip',
          "oTableTools": {
          "sSwfPath": "../media/swf/copy_csv_xls_pdf.swf",
          "aButtons": [
                "copy",
                "csv"
                ]
          },
      "aoColumns": [ 
      null,
  		{"bVisible": false},
      {"bVisible": false},
		  {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"sType": "scientific"},
      {"bVisible": false},
      null,
      null
		]          
    });
    });
</script>

******

## Fusion Gene Candidates

Fusion gene candidated are assessed using [Fusioncatcher](https://code.google.com/p/fusioncatcher/) `r citep("10.1186/gb-2011-12-1-r6")`. The oncogenicc potential of the fusion candidates is predicted using [OncoFuse](http://www.unav.es/genetica/oncofuse.html) `r citep("10.1093/bioinformatics/btt445")`.

```{r fusionCatcher, echo=FALSE, warning=FALSE, results='asis'}
```

<script type="text/javascript" charset="utf-8">
    $(document).ready(function() {
    var stdTable1 = $('#fusion_table').dataTable({
    "sDom": 'T<"clear">lfrtip',
    "oTableTools": {
          "sSwfPath": "../media/swf/copy_csv_xls_pdf.swf",
          "aButtons": [
                "copy",
                "csv"
                ]
    },
    "aoColumns": [ 
  		null,
			null,
      null,
		  {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      {"bVisible": false},
      null,
      null      
		], 
    //"iDisplayLength": 10,
    //"bPaginate": true,
    //"bProcessing": true,
    //"sScrollXInner": "110%",
    });
    
    //var tableId = 'fusion_table';
    //$('<div style="width: 100%; overflow: auto"></div>').append($('#' + tableId)).insertAfter($('#' + tableId + '_wrapper div').first())
    });
</script>

******

## SNPs

* **CHROM** The mutation’s chromosome.
* **POS** The mutation’s start position.
* **rsID**
* **Gene** The HUGO symbol of the gene.
* **Intogen_Driver**
* **SnpEff_Effect**
* **ClinVar_Rating**
* **ClinVar_ClinicalSignificance**
* **COSMIC_histology**
* **COSMIC_reference**
* **PharmGbk_Reaction**
* **PharmGbk_Drugs**
* **DrugBank_drug**
* **DrugBank_reaction**
* **DrugBank_reference**

```{r snpAnnotation, echo=FALSE, warning=FALSE, results='asis'}
```

<script type="text/javascript" charset="utf-8">
    $(document).ready(function() {
    var dogtable = $('#SNPs_table').dataTable({
        "sDom": 'T<"clear">lfrtip',
        "oTableTools": {
            "sSwfPath": "../media/swf/copy_csv_xls_pdf.swf",
            "aButtons": [
                "copy",
                "csv"
            ]
        },
        "aoColumns": [ 
          null,
          null,
          null,
          null,
          null,
          null,
          null,
          null,
          null,
          null,
          null,
          null,
          null,
          null,
          null,
          null,
          null
       ]         
      })
      });
</script>

******

# Drug matching based on overexpressed genes

```{r DoG, echo=FALSE, results='asis', message=FALSE}
```

<script type="text/javascript" charset="utf-8">
    $(document).ready(function() {
    var dogtable = $('#dog_table').dataTable({
        "sDom": 'T<"clear">lfrtip',
        "oTableTools": {
            "sSwfPath": "../media/swf/copy_csv_xls_pdf.swf",
            "aButtons": [
                "copy",
                "csv"
            ]
        }
      })
      });
</script>

******

# References 

```{r references, echo=FALSE, results='asis', warning=FALSE}
bibliography()

```


# Session Info

```{r, echo=FALSE}
sessionInfo()
date()
```
