Differential expresion using DESeq2
========================================================

<!-- dataTables -->
<script src="http://ajax.aspnetcdn.com/ajax/jquery.dataTables/1.9.4/jquery.dataTables.js"></script>
<link rel="stylesheet" href="http://ajax.aspnetcdn.com/ajax/jquery.dataTables/1.9.4/css/jquery.dataTables.css"/>
<script src="http://next.datatables.net/release-datatables/examples/resources/bootstrap/3/dataTables.bootstrap.js"></script>
<link rel="stylesheet" href="http://next.datatables.net/release-datatables/examples/resources/bootstrap/3/dataTables.bootstrap.css"/>

<!-- TableTools -->
<script src="http://cdn.aldu.net/jquery.datatables/extras/tabletools/2.0.3/js/TableTools.js"></script>
<link rel="stylesheet" href="http://cdn.aldu.net/jquery.datatables/extras/tabletools/2.0.3/css/TableTools.css"/>

<!-- ZeroClipboard -->
<script src="http://cdn.aldu.net/jquery.datatables/extras/tabletools/2.0.3/js/ZeroClipboard.js"></script>

<!-- ColViz -->
<script src="http://cdn.eventcore.com/js/plugins/dataTables.ColVis-1.07.js"></script>

## Load packages

```{r start}
library(DESeq2)
library(gplots)
library(RColorBrewer)

opts_chunk$set(warning = FALSE)
opts_chunk$set(message = FALSE)
opts_chunk$set(dev = 'pdf')
```

## Count matrix input

```{r htseq_count, results='asis'}
countFiles <- paste(config$HTSEQ_RESULTS, "/", dir(config$HTSEQ_RESULTS), sep='')
countNames <- gsub('_counts.txt', '', dir(config$HTSEQ_RESULTS))

countsDf <- NULL
for (i in countFiles) {
  dat <- read.csv(i, header=F, sep="\t", na.strings="", as.is=T)
  countsDf <- cbind(countsDf, dat[,2])
}
x1 <- dim(countsDf)[1]-4
x2 <- dim(countsDf)[1]
countsDf <- countsDf[-c(x1:x2),] # remove the last 5 lines, they hold no genes

rownames(countsDf) <- read.csv(i, header=F, sep="\t", na.strings="", as.is=T)$V1[-c(x1:x2)]
colnames(countsDf) <- countNames
write.csv(countsDf, sprintf("%s/counts_ALL.csv", config$DESEQ_RESULTS))

kable(head(countsDf), format='html', table.attr = 'id=\"count_table\"')
```

<script type="text/javascript" charset="utf-8">
    $(document).ready( function () {
    $('#count_table').dataTable({
          "scrollX": true,
          "sDom": 'T<"clear">lfrtip',
          "oTableTools": {
          "sSwfPath": "../media/swf/copy_csv_xls_pdf.swf",
          "aButtons": [
                "copy",
                "csv"
                ]
          }
    });
    });
</script>

*******

## Read in meta data

```{r meta_data, results='asis'}
meta <- read.csv2(config$DESEQ_META, header=T, stringsAsFactor=FALSE, sep=',')
meta <- meta[match(colnames(countsDf), meta$Sample), ] # sorte meta data by colnames of count matrix
rownames(meta) <- meta$Sample
meta$condition<-as.factor(meta$condition)
if (exists("config$DESEQ_REF")) { meta$condition<-relevel(meta$condition, ref=config$DESEQ_REF) }

kable(meta, format='html', table.attr = 'id=\"meta_table\"', row.names=FALSE)
```

<script type="text/javascript" charset="utf-8">
    $(document).ready( function () {
    $('#meta_table').dataTable({
          "sDom": 'T<"clear">lfrtip',
          "oTableTools": {
          "sSwfPath": "../media/swf/copy_csv_xls_pdf.swf",
          "aButtons": [
                "copy",
                "csv"
                ]
          }
    });
    });
</script>

*******

## dds object

```{r dds_object}
dds <- DESeqDataSetFromMatrix(countData = countsDf, 
                              colData = meta,
                              design = eval(parse(text=paste(config$DESIGN))))

print(dds)
```


## Differential Expression

```{r diff_expression}
dds <- DESeq(dds)
```


## Dispersion

```{r dispersion}
plotDispEsts(dds)
```


## MA Plot

```{r MAplot}
plotMA(dds,main="DESeq2",alpha=config$PVAL)
```

## Dealing with count outliers

```{r outliers}
ddsClean <- replaceOutliersWithTrimmedMean(dds)
ddsClean <- DESeq(ddsClean)
tab <- table(initial = results(dds)$padj < config$PVAL, cleaned = results(ddsClean)$padj < config$PVAL)
addmargins(tab)
```

## Differential expressed genes

```{r diffTable, results='asis'}
res <- results(ddsClean)
res <- res[order(res$padj),]
write.csv(res, sprintf("%s/DE_genes.csv", config$DESEQ_RESULTS))

kable(head(res, 100), format='html', table.attr = 'id=\"diff_table\"')

```

<script type="text/javascript" charset="utf-8">
    $(document).ready( function () {
    $('#diff_table').dataTable({
          "sDom": 'T<"clear">lfrtip',
          "oTableTools": {
          "sSwfPath": "../media/swf/copy_csv_xls_pdf.swf",
          "aButtons": [
                "copy",
                "csv"
                ]
          }
    });
    });
</script>

*******

## rlog Transformation for visualization

```{r transformation}
rld <- rlogTransformation(ddsClean, blind=TRUE)
```

## Heatmap of the count table

```{r heatmap_count_table}
select <- order(rowMeans(counts(dds,normalized=TRUE)),decreasing=TRUE)[1:30]
hmcol <- colorRampPalette(brewer.pal(9, "GnBu"))(100)

heatmap.2(assay(rld)[select,], 
          col = hmcol,
          Rowv = FALSE, 
          Colv = FALSE, 
          scale="none",
          dendrogram="none", 
          trace="none", 
          margin=c(10, 6))
```

## Sample to sample Distance

```{r heat_sample_to_sample_dist}
distsRL <- dist(t(assay(rld)))
mat <- as.matrix(distsRL)
rownames(mat) <- colnames(mat) <- with(colData(dds), paste(condition, libType, sep=" : "))
heatmap.2(mat, trace="none", col = rev(hmcol), margin=c(13, 13))
```

## PCA

```{r PCA}
print(plotPCA(rld, intgroup=c("condition", "libType")))
```






